---
title: "Guiding supervisors in AI forecasting"
output: html_notebook
---
## Experiment 2

```{r}
# reading data- experiment 2- part T1 and T2
library(readr)
exp2_t12 <- read.csv("experiment2_T1T2.csv")
nrow(exp2_t12)
ncol(exp2_t12)
head(exp2_t12)
```

```{r}
library(dplyr)
exp2_t12 <- filter(exp2_t12, test_round==0)
nrow(exp2_t12)
```

```{r}
exp2_t12$participant_id <- as.factor(exp2_t12$participant_id)
exp2_t12$code <- as.factor(exp2_t12$code)

exp2_t12$PQ1 <- as.factor(exp2_t12$PQ1)
exp2_t12$PQ2 <- as.factor(exp2_t12$PQ2)
exp2_t12$PQ3 <- as.factor(exp2_t12$PQ3)
exp2_t12$PQ4 <- as.factor(exp2_t12$PQ4)
exp2_t12$PQ6 <- as.factor(exp2_t12$PQ6)
exp2_t12$PQ7 <- as.factor(exp2_t12$PQ7)
exp2_t12$PQ8 <- as.factor(exp2_t12$PQ8)
exp2_t12$PQ9 <- as.factor(exp2_t12$PQ9)
exp2_t12$PQ10 <- as.factor(exp2_t12$PQ10)

exp2_t12$comp_skill <- as.factor(exp2_t12$comp_skill)
exp2_t12$forecasting_skill <- as.factor(exp2_t12$forecasting_skill)

exp2_t12$page <- as.factor(exp2_t12$page)
exp2_t12$round <- as.factor(exp2_t12$round)

exp2_t12$select_human <- as.factor(exp2_t12$select_human)

exp2_t12$pos_dir <- as.factor(exp2_t12$pos_dir)
exp2_t12$large_mag <- as.factor(exp2_t12$large_mag)
exp2_t12$sort_num <- as.factor(exp2_t12$sort_num)
exp2_t12$value <- as.factor(exp2_t12$value)
exp2_t12$h_AI_samedev <- as.factor(exp2_t12$h_AI_samedev)
exp2_t12$closer_to_mean <- as.factor(exp2_t12$closer_to_mean)

exp2_t12$earlier_rounds <- as.factor(exp2_t12$earlier_rounds)
exp2_t12$later_rounds <- as.factor(exp2_t12$later_rounds)
exp2_t12$directly_after_feedback <- as.factor(exp2_t12$directly_after_feedback)
exp2_t12$h_closer_tomean <- as.factor(exp2_t12$h_closer_tomean)

exp2_t12$guidance_followed_cat <- as.factor(exp2_t12$guidance_followed)
```

```{r}
prop.table(table(exp2_t12$select_human))
```

```{r}
# experiment 2- Part T1 and T2
prop.table(table(exp2_t12$gender))*100
prop.table(table(exp2_t12$education))*100
summary(exp2_t12$age)
prop.table(table(exp2_t12$work_status))*100
prop.table(table(exp2_t12$comp_skill))*100
prop.table(table(exp2_t12$forecasting_skill))*100
prop.table(table(exp2_t12$similar_exp))*100
```


```{r}
exp2_t12_pay <- exp2_t12 %>% group_by(participant_id) %>% summarize(payoff_tpp= sum(payoff)/48)

mean(exp2_t12_pay$payoff_tpp)
sd(exp2_t12_pay$payoff_tpp)
```

```{r}
tr1_weak <- filter(exp2_t12, tr== 60)
tr2_strong <- filter(exp2_t12, tr== 70)
```

```{r}
tr1_weak_pay <- tr1_weak %>% group_by(participant_id) %>% summarize(payoff_tpp= sum(payoff)/48)

mean(tr1_weak_pay$payoff_tpp)
sd(tr1_weak_pay$payoff_tpp)
```

```{r}
shapiro.test(tr1_weak_pay$payoff_tpp)
```

```{r}
tr2_strong_pay <- tr2_strong %>% group_by(participant_id) %>% summarize(payoff_tpp= sum(payoff)/48)

mean(tr2_strong_pay$payoff_tpp)
sd(tr2_strong_pay$payoff_tpp)
```

```{r}
shapiro.test(tr2_strong_pay$payoff_tpp)
```

```{r}
wilcox.test(tr1_weak_pay$payoff_tpp, tr2_strong_pay$payoff_tpp)
```

in line with the second paragraph section 4-6

## logistinc regression

### treatment 1

```{r}
tr1_weak_m <- select(tr1_weak, c("participant_id", "PQ1", "PQ2", "PQ3", "PQ4", "PQ6", "PQ7", "PQ8", "PQ9", "PQ10", "age", "gender", "education", "work_status", "comp_skill", "forecasting_skill", "similar_exp", "tinst_ttot", "select_human", "pos_dir", "large_mag", "closer_to_mean", "adj_vs_mean", "adj_vs_AI", "earlier_rounds", "later_rounds", "directly_after_feedback", "h_closer_tomean", "guidance_followed", "guidance_followed_cat"))
```


```{r}
model0_tr1_weak <- glm(select_human~ h_closer_tomean + adj_vs_AI+ earlier_rounds+ later_rounds+ directly_after_feedback+ large_mag* pos_dir, data=tr1_weak_m, family= binomial(link="logit"))
summary(model0_tr1_weak)
```

### treatment 2
```{r}
tr2_strong_m <- select(tr2_strong, c("participant_id", "PQ1", "PQ2", "PQ3", "PQ4", "PQ6", "PQ7", "PQ8", "PQ9", "PQ10", "age", "gender", "education", "work_status", "comp_skill", "forecasting_skill", "similar_exp", "tinst_ttot", "select_human", "pos_dir", "large_mag", "closer_to_mean", "adj_vs_mean", "adj_vs_AI", "earlier_rounds", "later_rounds", "directly_after_feedback", "h_closer_tomean", "guidance_followed"))
```

```{r}
model0_tr2_strong <- glm(select_human~ h_closer_tomean + adj_vs_AI+ earlier_rounds+ later_rounds+ directly_after_feedback+ large_mag* pos_dir, data=tr2_strong_m, family= binomial(link="logit"))
summary(model0_tr2_strong)
```

```{r}
# reading data- experiment 2- part T3 and T4
library(readr)
exp2_t34 <- read.csv("experiment2_T3T4.csv")
nrow(exp2_t34)
ncol(exp2_t34)
head(exp2_t34)
```

```{r}
library(dplyr)
exp2_t34 <- filter(exp2_t34, test_round==0)
nrow(exp2_t34)
```

```{r}
exp2_t34$participant_id <- as.factor(exp2_t34$participant_id)
exp2_t34$code <- as.factor(exp2_t34$code)

exp2_t34$PQ1 <- as.factor(exp2_t34$PQ1)
exp2_t34$PQ2 <- as.factor(exp2_t34$PQ2)
exp2_t34$PQ3 <- as.factor(exp2_t34$PQ3)
exp2_t34$PQ4 <- as.factor(exp2_t34$PQ4)
exp2_t34$PQ6 <- as.factor(exp2_t34$PQ6)
exp2_t34$PQ7 <- as.factor(exp2_t34$PQ7)
exp2_t34$PQ8 <- as.factor(exp2_t34$PQ8)
exp2_t34$PQ9 <- as.factor(exp2_t34$PQ9)
exp2_t34$PQ10 <- as.factor(exp2_t34$PQ10)

exp2_t34$comp_skill <- as.factor(exp2_t34$comp_skill)
exp2_t34$forecasting_skill <- as.factor(exp2_t34$forecasting_skill)

exp2_t34$page <- as.factor(exp2_t34$page)
exp2_t34$round <- as.factor(exp2_t34$round)

exp2_t34$select_human <- as.factor(exp2_t34$select_human)

exp2_t34$pos_dir <- as.factor(exp2_t34$pos_dir)
exp2_t34$large_mag <- as.factor(exp2_t34$large_mag)
exp2_t34$sort_num <- as.factor(exp2_t34$sort_num)
exp2_t34$value <- as.factor(exp2_t34$value)
exp2_t34$h_AI_samedev <- as.factor(exp2_t34$h_AI_samedev)
exp2_t34$closer_to_mean <- as.factor(exp2_t34$closer_to_mean)

exp2_t34$earlier_rounds <- as.factor(exp2_t34$earlier_rounds)
exp2_t34$later_rounds <- as.factor(exp2_t34$later_rounds)
exp2_t34$directly_after_feedback <- as.factor(exp2_t34$directly_after_feedback)
exp2_t34$h_closer_tomean <- as.factor(exp2_t34$h_closer_tomean)

exp2_t34$guidance_followed_cat <- as.factor(exp2_t34$guidance_followed)
```


```{r}
prop.table(table(exp2_t34$select_human))
```

```{r}
# experiment 2- Part T1 and T2
prop.table(table(exp2_t34$gender))*100
prop.table(table(exp2_t34$education))*100
summary(exp2_t12$age)
prop.table(table(exp2_t34$work_status))*100
prop.table(table(exp2_t34$comp_skill))*100
prop.table(table(exp2_t34$forecasting_skill))*100
prop.table(table(exp2_t34$similar_exp))*100
```


```{r}
exp2_t34_pay <- exp2_t34 %>% group_by(participant_id) %>% summarize(payoff_tpp= sum(payoff)/48)

mean(exp2_t34_pay$payoff_tpp)
sd(exp2_t34_pay$payoff_tpp)
```

```{r}
tr3_weak <- filter(exp2_t34, tr== 60)
tr4_strong <- filter(exp2_t34, tr== 70)
```

```{r}
tr3_weak_pay <- tr3_weak %>% group_by(participant_id) %>% summarize(payoff_tpp= sum(payoff)/48)

mean(tr3_weak_pay$payoff_tpp)
sd(tr3_weak_pay$payoff_tpp)
```

```{r}
shapiro.test(tr3_weak_pay$payoff_tpp)
```

```{r}
tr4_strong_pay <- tr4_strong %>% group_by(participant_id) %>% summarize(payoff_tpp= sum(payoff)/48)

mean(tr4_strong_pay$payoff_tpp)
sd(tr4_strong_pay$payoff_tpp)
```

```{r}
shapiro.test(tr4_strong_pay$payoff_tpp)
```

```{r}
wilcox.test(tr3_weak_pay$payoff_tpp, tr4_strong_pay$payoff_tpp)
```

in line with the second paragraph section 4-6

## logistinc regression

### treatment 3
```{r}
tr3_weak_m <- select(tr3_weak, c("participant_id", "PQ1", "PQ2", "PQ3", "PQ4", "PQ6", "PQ7", "PQ8", "PQ9", "PQ10", "age", "gender", "education", "work_status", "comp_skill", "forecasting_skill", "similar_exp", "tinst_ttot", "select_human", "pos_dir", "large_mag", "closer_to_mean", "adj_vs_mean", "adj_vs_AI", "earlier_rounds", "later_rounds", "directly_after_feedback", "h_closer_tomean", "guidance_followed", "guidance_followed_cat"))
```


```{r}
model0_tr3_weak <- glm(select_human~ h_closer_tomean + adj_vs_AI+ earlier_rounds+ later_rounds+ directly_after_feedback+ large_mag* pos_dir, data=tr3_weak_m, family= binomial(link="logit"))
summary(model0_tr3_weak)
```

### treatment 4
```{r}
tr4_strong_m <- select(tr4_strong, c("participant_id", "PQ1", "PQ2", "PQ3", "PQ4", "PQ6", "PQ7", "PQ8", "PQ9", "PQ10", "age", "gender", "education", "work_status", "comp_skill", "forecasting_skill", "similar_exp", "tinst_ttot", "select_human", "pos_dir", "large_mag", "closer_to_mean", "adj_vs_mean", "adj_vs_AI", "earlier_rounds", "later_rounds", "directly_after_feedback", "h_closer_tomean", "guidance_followed"))
```

```{r}
model0_tr4_strong <- glm(select_human~ h_closer_tomean + adj_vs_AI+ earlier_rounds+ later_rounds+ directly_after_feedback+ large_mag* pos_dir, data=tr4_strong_m, family= binomial(link="logit"))
summary(model0_tr4_strong)
```

### average guidance followed- statistical tests
```{r}
# focusing on downward adjusted cases
tr1_weak_neg <- filter(tr1_weak, pos_dir==0)
tr2_strong_neg <- filter(tr2_strong, pos_dir==0)
tr3_weak_neg <- filter(tr3_weak, pos_dir==0)
tr4_strong_neg <- filter(tr4_strong, pos_dir==0)
```

```{r}
tr1_weak_neg_guidancefollowed <- tr1_weak_neg %>% group_by(participant_id) %>% summarize(avg_guidance_followed= sum(guidance_followed==1)/24)

mean(tr1_weak_neg_guidancefollowed$avg_guidance_followed)
sd(tr1_weak_neg_guidancefollowed$avg_guidance_followed)
```

```{r}
tr2_strong_neg_guidancefollowed <- tr2_strong_neg %>% group_by(participant_id) %>% summarize(avg_guidance_followed= sum(guidance_followed==1)/24)

mean(tr2_strong_neg_guidancefollowed$avg_guidance_followed)
sd(tr2_strong_neg_guidancefollowed$avg_guidance_followed)
```

```{r}
tr3_weak_neg_guidancefollowed <- tr3_weak_neg %>% group_by(participant_id) %>% summarize(avg_guidance_followed= sum(guidance_followed==1)/24)

mean(tr3_weak_neg_guidancefollowed$avg_guidance_followed)
sd(tr3_weak_neg_guidancefollowed$avg_guidance_followed)
```

```{r}
tr4_strong_neg_guidancefollowed <- tr4_strong_neg %>% group_by(participant_id) %>% summarize(avg_guidance_followed= sum(guidance_followed==1)/24)

mean(tr4_strong_neg_guidancefollowed$avg_guidance_followed)
sd(tr4_strong_neg_guidancefollowed$avg_guidance_followed)
```

```{r}
shapiro.test(tr1_weak_neg_guidancefollowed$avg_guidance_followed)
shapiro.test(tr2_strong_neg_guidancefollowed$avg_guidance_followed)
shapiro.test(tr3_weak_neg_guidancefollowed$avg_guidance_followed)
shapiro.test(tr4_strong_neg_guidancefollowed$avg_guidance_followed)
```

```{r}
wilcox.test(tr1_weak_neg_guidancefollowed$avg_guidance_followed, tr2_strong_neg_guidancefollowed$avg_guidance_followed)
```

```{r}
wilcox.test(tr3_weak_neg_guidancefollowed$avg_guidance_followed, tr4_strong_neg_guidancefollowed$avg_guidance_followed)
```

There is no significant difference between average guidance followed in less and more salient treatments, when the level of detail is fixed.

```{r}
wilcox.test(tr1_weak_neg_guidancefollowed$avg_guidance_followed, tr3_weak_neg_guidancefollowed$avg_guidance_followed)
```

```{r}
wilcox.test(tr2_strong_neg_guidancefollowed$avg_guidance_followed, tr4_strong_neg_guidancefollowed$avg_guidance_followed)
```

There is a significant difference between less and more detailed treatments, when the guidance is more salient.

### MAE
```{r}
# MAE for human adjusted forecast, we presented the same forecasts to all subjects in different treatments, tr1_weak is used here as a sample
mean(abs(tr1_weak$demand - tr1_weak$human_forecast))
```

```{r}
# MAE for AI forecast, we presented the same forecasts to all subjects in different treatments, tr1_weak is used here as a sample
mean(abs(tr1_weak$demand - tr1_weak$AI_forecast))
```

To calculate MAE when guidance is fully followed in T1 and T2, we need to first understand the choices. Both T1 and T2 lead subjects to choose all human adjuted forecasts when there is a downward adjustment, and only AI forecast when there is an upward adjustment. This can be achieved easier in excel file of data.

```{r}
# mae for exp2- T1
tr1_weak_temp <- tr1_weak
tr1_weak_temp$ae <- abs(tr1_weak$selected_quant- tr1_weak$demand)
tr1_weak_mae <- tr1_weak_temp %>% group_by(participant_id) %>% summarize(mae= sum(ae)/48)
mean(tr1_weak_mae$mae)
sd(tr1_weak_mae$mae)
```

```{r}
# mae for exp2- T2
tr2_strong_temp <- tr2_strong
tr2_strong_temp$ae <- abs(tr2_strong$selected_quant- tr2_strong$demand)
tr2_strong_mae <- tr2_strong_temp %>% group_by(participant_id) %>% summarize(mae= sum(ae)/48)
mean(tr2_strong_mae$mae)
sd(tr2_strong_mae$mae)
```

```{r}
# mae for exp2- T1
tr3_weak_temp <- tr3_weak
tr3_weak_temp$ae <- abs(tr3_weak$selected_quant- tr3_weak$demand)
tr3_weak_mae <- tr3_weak_temp %>% group_by(participant_id) %>% summarize(mae= sum(ae)/48)
mean(tr3_weak_mae$mae)
sd(tr3_weak_mae$mae)
```

```{r}
# mae for exp2- T2
tr4_strong_temp <- tr4_strong
tr4_strong_temp$ae <- abs(tr4_strong$selected_quant- tr4_strong$demand)
tr4_strong_mae <- tr4_strong_temp %>% group_by(participant_id) %>% summarize(mae= sum(ae)/48)
mean(tr4_strong_mae$mae)
sd(tr4_strong_mae$mae)
```


